### 2023-02-27 Texfiles/Notes.tex
Add all code files
Adding files that I was previously working on in OneDrive.

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[authoryear]{natbib}
\usepackage{enumerate}
\setlength{\parindent}{0pt}
\usepackage[parfill]{parskip}
\date{Lent 2023}
\title{Dissertation notes}
\author{Tobias Leigh-Wood}
\usepackage[letterpaper, margin=1.1in]{geometry}
\linespread{1.1}

\begin{document}
\maketitle

\section*{To Do List}
\begin{itemize}
    \item What actually happened?
          Annuity provision peaked in 2012. Prior to the March 2011 budget individuals had to annuitise their pension pot.
          After 2011 they only had to annuitise up to a minimum income requirement (MIR) of £20,000. This minimum income kept most people using annuities.

          In March 2014 there was a surprise announcement that the MIR would be scrapped and individuals would be able to 'draw down' their pension pot instead.
          This meant the compulsory market for pension annuities basically stopped.

    \item When did the changes take place?
    \item When were the changes announced?
\end{itemize}

\end{document}