### 2023-02-28 Texfiles/Notes.tex
Create and practice with bibliography

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[authoryear]{natbib}
\usepackage{enumerate}
\setlength{\parindent}{0pt}
\usepackage[parfill]{parskip}
\date{Lent 2023}
\title{Dissertation notes}
\author{Tobias Leigh-Wood}
\usepackage[letterpaper, margin=1.1in]{geometry}
\linespread{1.1}

\begin{document}
\maketitle

\section*{To Do List}
\begin{itemize}
    \item What actually happened?
          Annuity provision peaked in 2012 (\cite{cannon_effect_2016}) Prior to the March 2011 budget individuals had to annuitise their pension pot.
          After 2011 they only had to annuitise up to a minimum income requirement (MIR) of £20,000. This minimum income kept most people using annuities.

          In March 2014 there was a surprise announcement that the MIR would be scrapped and individuals would be able to 'draw down' their pension pot instead.
          This meant the compulsory market for pension annuities basically stopped. Moreover, the amount needed to be classed as a 'small pot' increased meaning there were more pots that could be taken out all at once.
          In particular this reform meant the number of small annuities in the compulsory market was drastically reduced.

    \item When did the changes take place?
    \item When were the changes announced?
\end{itemize}

\bibliographystyle{chicago}
\bibliography{references}
\end{document}